---
title: "gwas_input_prep"
output: html_document
---


RUN = *Cmd+Shift+Enter*.  
ADD CHUNK = *Cmd+Option+I*.  
PREVIEW = *Cmd+Shift+K*.    

## Install Packages/Libaries
```{r}
# Install packages
#devtools::install_github("kenhanscombe/ukbtools", dependencies = TRUE)
#install.packages("Matching", dependencies=TRUE) 
#install.packages("MatchIt")
#install.packages("optmatch")

# Add libraries
library(ukbtools)
#library(optmatch)
#library(Matching)
#library(MatchIt)
```

## Load up data
```{r}
# Create data frames for second data release
my_ukb_data <- ukb_df("ukb40721")
dim(my_ukb_data)
```

## Define variables to keep in data frame
```{r}
# Define all necessary disease columns for IDC10. Arrays 0-65.
col_disease_icd10 <- c("diagnoses_icd10_f41270_0_0","diagnoses_icd10_f41270_0_1",
                       "diagnoses_icd10_f41270_0_2","diagnoses_icd10_f41270_0_3",
                       "diagnoses_icd10_f41270_0_4","diagnoses_icd10_f41270_0_5",
                       "diagnoses_icd10_f41270_0_6","diagnoses_icd10_f41270_0_7",
                       "diagnoses_icd10_f41270_0_8","diagnoses_icd10_f41270_0_9",
                       "diagnoses_icd10_f41270_0_10","diagnoses_icd10_f41270_0_11",
                       "diagnoses_icd10_f41270_0_12","diagnoses_icd10_f41270_0_13",
                       "diagnoses_icd10_f41270_0_14","diagnoses_icd10_f41270_0_15",
                       "diagnoses_icd10_f41270_0_16","diagnoses_icd10_f41270_0_17",
                       "diagnoses_icd10_f41270_0_18","diagnoses_icd10_f41270_0_19",
                       "diagnoses_icd10_f41270_0_20","diagnoses_icd10_f41270_0_21",
                       "diagnoses_icd10_f41270_0_22","diagnoses_icd10_f41270_0_23",
                       "diagnoses_icd10_f41270_0_24","diagnoses_icd10_f41270_0_25",
                       "diagnoses_icd10_f41270_0_26","diagnoses_icd10_f41270_0_27",
                       "diagnoses_icd10_f41270_0_28","diagnoses_icd10_f41270_0_29",
                       "diagnoses_icd10_f41270_0_30","diagnoses_icd10_f41270_0_31",
                       "diagnoses_icd10_f41270_0_32","diagnoses_icd10_f41270_0_33",
                       "diagnoses_icd10_f41270_0_34","diagnoses_icd10_f41270_0_35",
                       "diagnoses_icd10_f41270_0_36","diagnoses_icd10_f41270_0_37",
                       "diagnoses_icd10_f41270_0_38","diagnoses_icd10_f41270_0_39",
                       "diagnoses_icd10_f41270_0_40","diagnoses_icd10_f41270_0_41",
                       "diagnoses_icd10_f41270_0_42","diagnoses_icd10_f41270_0_43",
                       "diagnoses_icd10_f41270_0_44","diagnoses_icd10_f41270_0_45",
                       "diagnoses_icd10_f41270_0_46","diagnoses_icd10_f41270_0_47",
                       "diagnoses_icd10_f41270_0_48","diagnoses_icd10_f41270_0_49",
                       "diagnoses_icd10_f41270_0_50","diagnoses_icd10_f41270_0_51",
                       "diagnoses_icd10_f41270_0_52","diagnoses_icd10_f41270_0_53",
                       "diagnoses_icd10_f41270_0_54","diagnoses_icd10_f41270_0_55",
                       "diagnoses_icd10_f41270_0_56","diagnoses_icd10_f41270_0_57",
                       "diagnoses_icd10_f41270_0_58","diagnoses_icd10_f41270_0_59",
                       "diagnoses_icd10_f41270_0_60","diagnoses_icd10_f41270_0_61",
                       "diagnoses_icd10_f41270_0_62","diagnoses_icd10_f41270_0_63",
                       "diagnoses_icd10_f41270_0_64","diagnoses_icd10_f41270_0_65",
                       "diagnoses_icd10_f41270_0_66","diagnoses_icd10_f41270_0_67",
                       "diagnoses_icd10_f41270_0_68","diagnoses_icd10_f41270_0_69",
                       "diagnoses_icd10_f41270_0_70","diagnoses_icd10_f41270_0_71",
                       "diagnoses_icd10_f41270_0_72","diagnoses_icd10_f41270_0_73",
                       "diagnoses_icd10_f41270_0_74","diagnoses_icd10_f41270_0_75",
                       "diagnoses_icd10_f41270_0_76","diagnoses_icd10_f41270_0_77",
                       "diagnoses_icd10_f41270_0_78","diagnoses_icd10_f41270_0_79",
                       "diagnoses_icd10_f41270_0_80","diagnoses_icd10_f41270_0_81",
                       "diagnoses_icd10_f41270_0_82","diagnoses_icd10_f41270_0_83",
                       "diagnoses_icd10_f41270_0_84","diagnoses_icd10_f41270_0_85",
                       "diagnoses_icd10_f41270_0_86","diagnoses_icd10_f41270_0_87",
                       "diagnoses_icd10_f41270_0_88","diagnoses_icd10_f41270_0_89",
                       "diagnoses_icd10_f41270_0_90","diagnoses_icd10_f41270_0_91",
                       "diagnoses_icd10_f41270_0_92","diagnoses_icd10_f41270_0_93",
                       "diagnoses_icd10_f41270_0_94","diagnoses_icd10_f41270_0_95",
                       "diagnoses_icd10_f41270_0_96","diagnoses_icd10_f41270_0_97",
                       "diagnoses_icd10_f41270_0_98","diagnoses_icd10_f41270_0_99",
                       "diagnoses_icd10_f41270_0_100","diagnoses_icd10_f41270_0_101",
                       "diagnoses_icd10_f41270_0_102","diagnoses_icd10_f41270_0_103",
                       "diagnoses_icd10_f41270_0_104","diagnoses_icd10_f41270_0_105",
                       "diagnoses_icd10_f41270_0_106","diagnoses_icd10_f41270_0_107",
                       "diagnoses_icd10_f41270_0_108","diagnoses_icd10_f41270_0_109",
                       "diagnoses_icd10_f41270_0_110","diagnoses_icd10_f41270_0_111",
                       "diagnoses_icd10_f41270_0_112","diagnoses_icd10_f41270_0_113",
                       "diagnoses_icd10_f41270_0_114","diagnoses_icd10_f41270_0_115",
                       "diagnoses_icd10_f41270_0_116","diagnoses_icd10_f41270_0_117",
                       "diagnoses_icd10_f41270_0_118","diagnoses_icd10_f41270_0_119",
                       "diagnoses_icd10_f41270_0_120","diagnoses_icd10_f41270_0_121",
                       "diagnoses_icd10_f41270_0_122","diagnoses_icd10_f41270_0_123",
                       "diagnoses_icd10_f41270_0_124","diagnoses_icd10_f41270_0_125",
                       "diagnoses_icd10_f41270_0_126","diagnoses_icd10_f41270_0_127",
                       "diagnoses_icd10_f41270_0_128","diagnoses_icd10_f41270_0_129",
                       "diagnoses_icd10_f41270_0_130","diagnoses_icd10_f41270_0_131",
                       "diagnoses_icd10_f41270_0_132","diagnoses_icd10_f41270_0_133",
                       "diagnoses_icd10_f41270_0_134","diagnoses_icd10_f41270_0_135",
                       "diagnoses_icd10_f41270_0_136","diagnoses_icd10_f41270_0_137",
                       "diagnoses_icd10_f41270_0_138","diagnoses_icd10_f41270_0_139",
                       "diagnoses_icd10_f41270_0_140","diagnoses_icd10_f41270_0_141",
                       "diagnoses_icd10_f41270_0_142","diagnoses_icd10_f41270_0_143",
                       "diagnoses_icd10_f41270_0_144","diagnoses_icd10_f41270_0_145",
                       "diagnoses_icd10_f41270_0_146","diagnoses_icd10_f41270_0_147",
                       "diagnoses_icd10_f41270_0_148","diagnoses_icd10_f41270_0_149",
                       "diagnoses_icd10_f41270_0_150","diagnoses_icd10_f41270_0_151",
                       "diagnoses_icd10_f41270_0_152","diagnoses_icd10_f41270_0_153",
                       "diagnoses_icd10_f41270_0_154","diagnoses_icd10_f41270_0_155",
                       "diagnoses_icd10_f41270_0_156","diagnoses_icd10_f41270_0_157",
                       "diagnoses_icd10_f41270_0_158","diagnoses_icd10_f41270_0_159",
                       "diagnoses_icd10_f41270_0_160","diagnoses_icd10_f41270_0_161",
                       "diagnoses_icd10_f41270_0_162","diagnoses_icd10_f41270_0_163",
                       "diagnoses_icd10_f41270_0_164","diagnoses_icd10_f41270_0_165",
                       "diagnoses_icd10_f41270_0_166","diagnoses_icd10_f41270_0_167",
                       "diagnoses_icd10_f41270_0_168","diagnoses_icd10_f41270_0_169",
                       "diagnoses_icd10_f41270_0_170","diagnoses_icd10_f41270_0_171",
                       "diagnoses_icd10_f41270_0_172","diagnoses_icd10_f41270_0_173",
                       "diagnoses_icd10_f41270_0_174","diagnoses_icd10_f41270_0_175",
                       "diagnoses_icd10_f41270_0_176","diagnoses_icd10_f41270_0_177",
                       "diagnoses_icd10_f41270_0_178","diagnoses_icd10_f41270_0_179",
                       "diagnoses_icd10_f41270_0_180","diagnoses_icd10_f41270_0_181",
                       "diagnoses_icd10_f41270_0_182","diagnoses_icd10_f41270_0_183",
                       "diagnoses_icd10_f41270_0_184","diagnoses_icd10_f41270_0_185",
                       "diagnoses_icd10_f41270_0_186","diagnoses_icd10_f41270_0_187",
                       "diagnoses_icd10_f41270_0_188","diagnoses_icd10_f41270_0_189",
                       "diagnoses_icd10_f41270_0_190","diagnoses_icd10_f41270_0_191",
                       "diagnoses_icd10_f41270_0_192","diagnoses_icd10_f41270_0_193",
                       "diagnoses_icd10_f41270_0_194","diagnoses_icd10_f41270_0_195",
                       "diagnoses_icd10_f41270_0_196","diagnoses_icd10_f41270_0_197",
                       "diagnoses_icd10_f41270_0_198","diagnoses_icd10_f41270_0_199",
                       "diagnoses_icd10_f41270_0_200","diagnoses_icd10_f41270_0_201",
                       "diagnoses_icd10_f41270_0_202","diagnoses_icd10_f41270_0_203",
                       "diagnoses_icd10_f41270_0_204","diagnoses_icd10_f41270_0_205",
                       "diagnoses_icd10_f41270_0_206","diagnoses_icd10_f41270_0_207",
                       "diagnoses_icd10_f41270_0_208","diagnoses_icd10_f41270_0_209",
                       "diagnoses_icd10_f41270_0_210","diagnoses_icd10_f41270_0_211",
                       "diagnoses_icd10_f41270_0_212")

# Additional variables needed for analysis
gwas_vars <- c("eid",
               "year_of_birth_f34_0_0","month_of_birth_f52_0_0",
               "sex_f31_0_0",
               "current_tobacco_smoking_f1239_0_0","current_tobacco_smoking_f1239_1_0",
               "current_tobacco_smoking_f1239_2_0","current_tobacco_smoking_f1239_3_0",
               "ethnic_background_f21000_0_0",
               "ethnic_background_f21000_1_0",
               "ethnic_background_f21000_2_0",
               "body_mass_index_bmi_f21001_0_0")
```

## Only keep needed variables
```{r}
my_ukb_data <- subset(my_ukb_data, select=c(col_disease_icd10,gwas_vars))
```

## Add extra variables to dataframe as needed
```{r}
# Add birthday variable, in R date format
my_ukb_data$birthday <- as.Date(sprintf("%d-%s-%d", my_ukb_data$year_of_birth_f34_0_0, 
                                        my_ukb_data$month_of_birth_f52_0_0,1), format="%Y-%B-%d")
# Add age
my_ukb_data$age_years <- as.numeric(floor((Sys.Date() - my_ukb_data$birthday)/365.25))
```

# Extract bad data point
```{r}
my_ukb_data <- my_ukb_data[-c(223169),]
```


## Make a UKB data-field to variable name table for reference or lookup.
```{r}
#my_ukb_key <- ukb_df_field("ukb40721")
#my_ukb_key
```

# Extract Caucasian patients
```{r}
table(my_ukb_data$ethnic_background_f21000_0_0)
#white <- with(my_ukb_data,
#                  which(ethnic_background_f21000_0_0 %in%
#                       c("White","Any other white background","British","Irish")))
white <- with(my_ukb_data,
                  which(ethnic_background_f21000_0_0 %in%
                       c("White","Any other white background","British",
                         "Irish", "White and Black Caribbean",
                         "White and Black African", "White and Asian")))
dim(my_ukb_data)[1]
length(white)
my_ukb_data <- my_ukb_data[white,]
```

## Define constants
```{r}
num_diagnosis_icd10 <- 410318
num_cancer_icd10 <- 83480
num_diagnosis_self <- 375036
num_cancer_self <- 41697

# Create list of all possible indicies
num_subjects <- nrow(my_ukb_data)
num_var <- ncol(my_ukb_data)
all_indices <- seq(1, nrow(my_ukb_data), 1)
```

## Grab data frame indices
```{r}
# CF ICD10
# Extract patients with K850-K859 acute pancreatitis designation via ICD 10
cf_conditions_icd10 <- c("E840","E841","E8411","E8419","E848", "E849")
cf_icd10_e84x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", cf_conditions_icd10))>0)
paste0("# with ICD10 AP ", length(cf_icd10_e84x))

# AP ICD10
# Extract patients with K850-K859 acute pancreatitis designation via ICD 10
ap_conditions_icd10 <- c("K85","K850","K851","K852","K853", "K858", "K859")
ap_icd10_k85x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", ap_conditions_icd10))>0)
paste0("# with ICD10 AP ", length(ap_icd10_k85x))

# CP ICD10
# Extract patients with K860-861 chronic pancreatitis designation via ICD 10
cp_conditions_icd10 <- c("K860","K861")
cp_icd10_k86x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", cp_conditions_icd10))>0)
paste0("# with ICD10 CP ", length(cp_icd10_k86x))

# AP & CP ICD10
panc_idc10_ix <- sort(union(ap_icd10_k85x, cp_icd10_k86x))
paste0("# with ICD10 AP or CP ", length(panc_idc10_ix))
paste0("# with both ICD10 AP and CP ", length(intersect(ap_icd10_k85x, cp_icd10_k86x)))

# CONTROLS ICD10
# Extract patients WITHOUT ICD10 AP or CP designation
control_icd10_total <- setdiff(all_indices, panc_idc10_ix)
paste0("# controls ", length(control_icd10_total))

# Grab samples of control population
#mult_factor <- 12 # Memory issues
#mult_factor <- 9
mult_factor <- 6
s <- length(panc_idc10_ix) * mult_factor
#s <- 5000
set.seed(5750)
#control_icd10_sample500 <- sample(control_icd10_total, s, replace = FALSE, prob = NULL)
#s <- 2000
#control_icd10_sample2000 <- sample(control_icd10_total, s, replace = FALSE, prob = NULL)
control_icd10_sample <- sample(control_icd10_total, s, replace = FALSE, prob = NULL)

# Test for correctness
(nrow(my_ukb_data) - length(control_icd10_total)) == (length(ap_icd10_k85x) + 
                                                     length(cp_icd10_k86x) - length(intersect(ap_icd10_k85x,cp_icd10_k86x)))
```

## Subset patient populations
```{r}
# Subset AP ICD10 population
#ap_icd10_data <- my_ukb_data[ap_icd10_k85x,]
#dim(ap_icd10_data)

# Subset CP ICD10 population
#cp_icd10_data <- my_ukb_data[cp_icd10_k86x,]
#dim(cp_icd10_data)

# Subset AP/CP ICD10 population
#panc_icd10_data <- my_ukb_data[panc_idc10_ix,]
#dim(panc_icd10_data)

# Subset ICD10 control sample population
#control_icd10_data <- my_ukb_data[control_icd10_sample2000,]
#control_icd10_data <- my_ukb_data[control_icd10_sample,]
#dim(control_icd10_data)
```



# Create TEXTFILE_OF_4K_IIDS.txt
```{r}
eids <- my_ukb_data$eid
x <- cbind(eids[panc_idc10_ix], eids[panc_idc10_ix])
#y <- cbind(eids[control_icd10_sample2000], eids[control_icd10_sample2000])
y <- cbind(eids[control_icd10_sample], eids[control_icd10_sample])
z <- rbind(x,y)
#fname <- "case_control_ids.txt"

fname <- "case_control_ids_6x.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

fname <- "case_only_ids_6x.txt"
write.table(x, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

fname <- "control_only_ids_6x.txt"
write.table(y, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

```

# Create pancreatitis_vs_control_phenotypes.txt
```{r}
x <- cbind(eids[panc_idc10_ix], eids[panc_idc10_ix],1)
#y <- cbind(eids[control_icd10_sample2000], eids[control_icd10_sample2000],0)
y <- cbind(eids[control_icd10_sample], eids[control_icd10_sample],0)
z <- rbind(x,y)
#fname <- "pancreatitis_vs_control_phenotypes.txt"
fname <- "pancreatitis_vs_control_phenotypes_6x.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

# Create pancreatitis_vs_control_discrete_covars.txt
```{r}

# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)
table(sex.c)
table(smoking.0c)
table(smoking.1c)
table(smoking.2c)
table(smoking.3c)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
            (
              (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
              (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
              (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
              (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
            ),
            "1",
            ifelse(
              (
                (smoking.0c == "No") |
                (smoking.1c == "No") | 
                (smoking.2c == "No") | 
                (smoking.3c == "No")
              ),
              "0",
              "NA"
            )
)

x <- cbind(eids[panc_idc10_ix], 
           eids[panc_idc10_ix],
           sex.c[panc_idc10_ix], 
           smoking[panc_idc10_ix])
#y <- cbind(eids[control_icd10_sample2000],
#           eids[control_icd10_sample2000],
#           sex.c[control_icd10_sample2000], 
#           smoking[control_icd10_sample2000])
y <- cbind(eids[control_icd10_sample],
           eids[control_icd10_sample],
           sex.c[control_icd10_sample], 
           smoking[control_icd10_sample])
z <- rbind(x,y)

# Discover NA indices for error checking
na.ix <- which(is.na(z[,4]))
#fname <- "pancreatitis_vs_control_discrete_covars.txt"
fname <- "pancreatitis_vs_control_discrete_covars_6x.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

# Create pancreatitis_vs_control_quantitative_covars.txt 
```{r}
# Extract BMI, all other quantititative covars are PCs which will be added later

# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0

# TODO(DMS): Alternately, can do an avg of intances 0-3??

# Look at BMI avgs
summary(bmi)

x <- cbind(eids[panc_idc10_ix], 
           eids[panc_idc10_ix],
           bmi[panc_idc10_ix])
#y <- cbind(eids[control_icd10_sample2000],
#           eids[control_icd10_sample2000],
#           bmi[control_icd10_sample2000])
y <- cbind(eids[control_icd10_sample],
           eids[control_icd10_sample],
           bmi[control_icd10_sample])
z <- rbind(x,y)
#fname <- "pancreatitis_vs_control_quantitative_covars_bmiOnly.txt"
fname <- "pancreatitis_vs_control_quantitative_covars_6x.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

```{r}
pca <- read.delim(file="ukb_cases.eigenvec", header=FALSE, sep=" ", dec=".")
pca$V2 <- NULL
#zeta <- as.data.frame(z)
zeta <- read.delim(file="pancreatitis_vs_control_quantitative_covars_bmiOnly_matched.txt", header=FALSE, sep=" ", dec=".")
zeta$ord <- 1:nrow(zeta) 
quant <- merge(zeta, pca, by.x="V1", by.y="V1", all.x=TRUE)
quant <- quant[order(quant$ord), ]
quant$ord <- NULL
quant <- as.matrix(quant)

fname <- "pancreatitis_vs_control_quantitative_covars.txt"
write.table(quant, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```


RUN = *Cmd+Shift+Enter*.  
ADD CHUNK = *Cmd+Option+I*.  
PREVIEW = *Cmd+Shift+K*.    

