---
title: "gwas_prep_matched"
author: "dspagnolo"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Install Packages/Libaries
```{r}
# Install packages
#devtools::install_github("kenhanscombe/ukbtools", dependencies = TRUE)
#install.packages("MatchIt")


# Add libraries
library(ukbtools)
library(MatchIt)
```

## Load up data
```{r}
# Create data frames for second data release
my_ukb_data <- ukb_df("ukb40721")
dim(my_ukb_data)
object.size(my_ukb_data)
```

## Define variables to keep in data frame
```{r}
# Define all necessary disease columns for IDC10. Arrays 0-65.
col_disease_icd10 <- c("diagnoses_icd10_f41270_0_0","diagnoses_icd10_f41270_0_1",
                       "diagnoses_icd10_f41270_0_2","diagnoses_icd10_f41270_0_3",
                       "diagnoses_icd10_f41270_0_4","diagnoses_icd10_f41270_0_5",
                       "diagnoses_icd10_f41270_0_6","diagnoses_icd10_f41270_0_7",
                       "diagnoses_icd10_f41270_0_8","diagnoses_icd10_f41270_0_9",
                       "diagnoses_icd10_f41270_0_10","diagnoses_icd10_f41270_0_11",
                       "diagnoses_icd10_f41270_0_12","diagnoses_icd10_f41270_0_13",
                       "diagnoses_icd10_f41270_0_14","diagnoses_icd10_f41270_0_15",
                       "diagnoses_icd10_f41270_0_16","diagnoses_icd10_f41270_0_17",
                       "diagnoses_icd10_f41270_0_18","diagnoses_icd10_f41270_0_19",
                       "diagnoses_icd10_f41270_0_20","diagnoses_icd10_f41270_0_21",
                       "diagnoses_icd10_f41270_0_22","diagnoses_icd10_f41270_0_23",
                       "diagnoses_icd10_f41270_0_24","diagnoses_icd10_f41270_0_25",
                       "diagnoses_icd10_f41270_0_26","diagnoses_icd10_f41270_0_27",
                       "diagnoses_icd10_f41270_0_28","diagnoses_icd10_f41270_0_29",
                       "diagnoses_icd10_f41270_0_30","diagnoses_icd10_f41270_0_31",
                       "diagnoses_icd10_f41270_0_32","diagnoses_icd10_f41270_0_33",
                       "diagnoses_icd10_f41270_0_34","diagnoses_icd10_f41270_0_35",
                       "diagnoses_icd10_f41270_0_36","diagnoses_icd10_f41270_0_37",
                       "diagnoses_icd10_f41270_0_38","diagnoses_icd10_f41270_0_39",
                       "diagnoses_icd10_f41270_0_40","diagnoses_icd10_f41270_0_41",
                       "diagnoses_icd10_f41270_0_42","diagnoses_icd10_f41270_0_43",
                       "diagnoses_icd10_f41270_0_44","diagnoses_icd10_f41270_0_45",
                       "diagnoses_icd10_f41270_0_46","diagnoses_icd10_f41270_0_47",
                       "diagnoses_icd10_f41270_0_48","diagnoses_icd10_f41270_0_49",
                       "diagnoses_icd10_f41270_0_50","diagnoses_icd10_f41270_0_51",
                       "diagnoses_icd10_f41270_0_52","diagnoses_icd10_f41270_0_53",
                       "diagnoses_icd10_f41270_0_54","diagnoses_icd10_f41270_0_55",
                       "diagnoses_icd10_f41270_0_56","diagnoses_icd10_f41270_0_57",
                       "diagnoses_icd10_f41270_0_58","diagnoses_icd10_f41270_0_59",
                       "diagnoses_icd10_f41270_0_60","diagnoses_icd10_f41270_0_61",
                       "diagnoses_icd10_f41270_0_62","diagnoses_icd10_f41270_0_63",
                       "diagnoses_icd10_f41270_0_64","diagnoses_icd10_f41270_0_65",
                       "diagnoses_icd10_f41270_0_66","diagnoses_icd10_f41270_0_67",
                       "diagnoses_icd10_f41270_0_68","diagnoses_icd10_f41270_0_69",
                       "diagnoses_icd10_f41270_0_70","diagnoses_icd10_f41270_0_71",
                       "diagnoses_icd10_f41270_0_72","diagnoses_icd10_f41270_0_73",
                       "diagnoses_icd10_f41270_0_74","diagnoses_icd10_f41270_0_75",
                       "diagnoses_icd10_f41270_0_76","diagnoses_icd10_f41270_0_77",
                       "diagnoses_icd10_f41270_0_78","diagnoses_icd10_f41270_0_79",
                       "diagnoses_icd10_f41270_0_80","diagnoses_icd10_f41270_0_81",
                       "diagnoses_icd10_f41270_0_82","diagnoses_icd10_f41270_0_83",
                       "diagnoses_icd10_f41270_0_84","diagnoses_icd10_f41270_0_85",
                       "diagnoses_icd10_f41270_0_86","diagnoses_icd10_f41270_0_87",
                       "diagnoses_icd10_f41270_0_88","diagnoses_icd10_f41270_0_89",
                       "diagnoses_icd10_f41270_0_90","diagnoses_icd10_f41270_0_91",
                       "diagnoses_icd10_f41270_0_92","diagnoses_icd10_f41270_0_93",
                       "diagnoses_icd10_f41270_0_94","diagnoses_icd10_f41270_0_95",
                       "diagnoses_icd10_f41270_0_96","diagnoses_icd10_f41270_0_97",
                       "diagnoses_icd10_f41270_0_98","diagnoses_icd10_f41270_0_99",
                       "diagnoses_icd10_f41270_0_100","diagnoses_icd10_f41270_0_101",
                       "diagnoses_icd10_f41270_0_102","diagnoses_icd10_f41270_0_103",
                       "diagnoses_icd10_f41270_0_104","diagnoses_icd10_f41270_0_105",
                       "diagnoses_icd10_f41270_0_106","diagnoses_icd10_f41270_0_107",
                       "diagnoses_icd10_f41270_0_108","diagnoses_icd10_f41270_0_109",
                       "diagnoses_icd10_f41270_0_110","diagnoses_icd10_f41270_0_111",
                       "diagnoses_icd10_f41270_0_112","diagnoses_icd10_f41270_0_113",
                       "diagnoses_icd10_f41270_0_114","diagnoses_icd10_f41270_0_115",
                       "diagnoses_icd10_f41270_0_116","diagnoses_icd10_f41270_0_117",
                       "diagnoses_icd10_f41270_0_118","diagnoses_icd10_f41270_0_119",
                       "diagnoses_icd10_f41270_0_120","diagnoses_icd10_f41270_0_121",
                       "diagnoses_icd10_f41270_0_122","diagnoses_icd10_f41270_0_123",
                       "diagnoses_icd10_f41270_0_124","diagnoses_icd10_f41270_0_125",
                       "diagnoses_icd10_f41270_0_126","diagnoses_icd10_f41270_0_127",
                       "diagnoses_icd10_f41270_0_128","diagnoses_icd10_f41270_0_129",
                       "diagnoses_icd10_f41270_0_130","diagnoses_icd10_f41270_0_131",
                       "diagnoses_icd10_f41270_0_132","diagnoses_icd10_f41270_0_133",
                       "diagnoses_icd10_f41270_0_134","diagnoses_icd10_f41270_0_135",
                       "diagnoses_icd10_f41270_0_136","diagnoses_icd10_f41270_0_137",
                       "diagnoses_icd10_f41270_0_138","diagnoses_icd10_f41270_0_139",
                       "diagnoses_icd10_f41270_0_140","diagnoses_icd10_f41270_0_141",
                       "diagnoses_icd10_f41270_0_142","diagnoses_icd10_f41270_0_143",
                       "diagnoses_icd10_f41270_0_144","diagnoses_icd10_f41270_0_145",
                       "diagnoses_icd10_f41270_0_146","diagnoses_icd10_f41270_0_147",
                       "diagnoses_icd10_f41270_0_148","diagnoses_icd10_f41270_0_149",
                       "diagnoses_icd10_f41270_0_150","diagnoses_icd10_f41270_0_151",
                       "diagnoses_icd10_f41270_0_152","diagnoses_icd10_f41270_0_153",
                       "diagnoses_icd10_f41270_0_154","diagnoses_icd10_f41270_0_155",
                       "diagnoses_icd10_f41270_0_156","diagnoses_icd10_f41270_0_157",
                       "diagnoses_icd10_f41270_0_158","diagnoses_icd10_f41270_0_159",
                       "diagnoses_icd10_f41270_0_160","diagnoses_icd10_f41270_0_161",
                       "diagnoses_icd10_f41270_0_162","diagnoses_icd10_f41270_0_163",
                       "diagnoses_icd10_f41270_0_164","diagnoses_icd10_f41270_0_165",
                       "diagnoses_icd10_f41270_0_166","diagnoses_icd10_f41270_0_167",
                       "diagnoses_icd10_f41270_0_168","diagnoses_icd10_f41270_0_169",
                       "diagnoses_icd10_f41270_0_170","diagnoses_icd10_f41270_0_171",
                       "diagnoses_icd10_f41270_0_172","diagnoses_icd10_f41270_0_173",
                       "diagnoses_icd10_f41270_0_174","diagnoses_icd10_f41270_0_175",
                       "diagnoses_icd10_f41270_0_176","diagnoses_icd10_f41270_0_177",
                       "diagnoses_icd10_f41270_0_178","diagnoses_icd10_f41270_0_179",
                       "diagnoses_icd10_f41270_0_180","diagnoses_icd10_f41270_0_181",
                       "diagnoses_icd10_f41270_0_182","diagnoses_icd10_f41270_0_183",
                       "diagnoses_icd10_f41270_0_184","diagnoses_icd10_f41270_0_185",
                       "diagnoses_icd10_f41270_0_186","diagnoses_icd10_f41270_0_187",
                       "diagnoses_icd10_f41270_0_188","diagnoses_icd10_f41270_0_189",
                       "diagnoses_icd10_f41270_0_190","diagnoses_icd10_f41270_0_191",
                       "diagnoses_icd10_f41270_0_192","diagnoses_icd10_f41270_0_193",
                       "diagnoses_icd10_f41270_0_194","diagnoses_icd10_f41270_0_195",
                       "diagnoses_icd10_f41270_0_196","diagnoses_icd10_f41270_0_197",
                       "diagnoses_icd10_f41270_0_198","diagnoses_icd10_f41270_0_199",
                       "diagnoses_icd10_f41270_0_200","diagnoses_icd10_f41270_0_201",
                       "diagnoses_icd10_f41270_0_202","diagnoses_icd10_f41270_0_203",
                       "diagnoses_icd10_f41270_0_204","diagnoses_icd10_f41270_0_205",
                       "diagnoses_icd10_f41270_0_206","diagnoses_icd10_f41270_0_207",
                       "diagnoses_icd10_f41270_0_208","diagnoses_icd10_f41270_0_209",
                       "diagnoses_icd10_f41270_0_210","diagnoses_icd10_f41270_0_211",
                       "diagnoses_icd10_f41270_0_212")

# Additional variables needed for analysis
gwas_vars <- c("eid",
               "year_of_birth_f34_0_0","month_of_birth_f52_0_0",
               "sex_f31_0_0",
               "current_tobacco_smoking_f1239_0_0","current_tobacco_smoking_f1239_1_0",
               "current_tobacco_smoking_f1239_2_0","current_tobacco_smoking_f1239_3_0",
               "body_mass_index_bmi_f21001_0_0")
```

## Only keep needed variables
```{r}
my_ukb_data <- subset(my_ukb_data, select=c(col_disease_icd10,gwas_vars))
dim(my_ukb_data)
object.size(my_ukb_data)
```

## Add extra variables to dataframe as needed
```{r}
# Add birthday variable, in R date format
my_ukb_data$birthday <- as.Date(sprintf("%d-%s-%d", my_ukb_data$year_of_birth_f34_0_0, 
                                        my_ukb_data$month_of_birth_f52_0_0,1), format="%Y-%B-%d")
# Add age
my_ukb_data$age_years <- as.numeric(floor((Sys.Date() - my_ukb_data$birthday)/365.25))
```

# Extract bad data point
```{r}
my_ukb_data <- my_ukb_data[-c(223169),]
```

## Grab data frame indices
```{r}
# AP ICD10
# Extract patients with K850-K859 acute pancreatitis designation via ICD 10
ap_conditions_icd10 <- c("K85","K850","K851","K852","K853", "K858", "K859")
ap_icd10_k85x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", ap_conditions_icd10))>0)
paste0("# with ICD10 AP ", length(ap_icd10_k85x))

# CP ICD10
# Extract patients with K860-861 chronic pancreatitis designation via ICD 10
cp_conditions_icd10 <- c("K860","K861")
cp_icd10_k86x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", cp_conditions_icd10))>0)
paste0("# with ICD10 CP ", length(cp_icd10_k86x))

# AP & CP ICD10
panc_idc10_ix <- sort(union(ap_icd10_k85x, cp_icd10_k86x))
paste0("# with ICD10 AP or CP ", length(panc_idc10_ix))
paste0("# with both ICD10 AP and CP ", length(intersect(ap_icd10_k85x, cp_icd10_k86x)))
```

# IF looking at just AP or CP, remove CP-only cases or AP-only cases so they don't end up as controls
```{r}
# Find indices
ap_cp <- intersect(ap_icd10_k85x, cp_icd10_k86x)
ap_only <- setdiff(ap_icd10_k85x, ap_cp)
cp_only <- setdiff(cp_icd10_k86x, ap_cp)

# Remove AP-only patients for CP study. Recompute indices
# my_ukb_data <- my_ukb_data[-ap_only,]
# cp_icd10_k86x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", cp_conditions_icd10))>0)

# Remove CP patients (incl AP-CP) for AP study
my_ukb_data <- my_ukb_data[-cp_icd10_k86x,]
ap_icd10_k85x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", ap_conditions_icd10))>0)
```


## Create minimal data fram for MatchIt. MatchIt requires no NA values in ANY column
```{r}
# Create "treatment" vector to identify observations in the "treatment regime"
num_subjects <- nrow(my_ukb_data)
case_vec <- logical(num_subjects)
#case_vec[panc_idc10_ix] <- TRUE # FOR AP & CP
#case_vec[cp_icd10_k86x] <- TRUE # FOR CP study
case_vec[ap_icd10_k85x] <- TRUE # FOR AP study

# Add treatment to data fram
#my_ukb_data$treat <- case_vec
my_ukb_data$treat <- as.integer(case_vec)

# Extract minimal data frame w/ variables needed for analysis
min_vars <- c("eid",
              "age_years",
              "sex_f31_0_0",
              "current_tobacco_smoking_f1239_0_0","current_tobacco_smoking_f1239_1_0",
              "current_tobacco_smoking_f1239_2_0","current_tobacco_smoking_f1239_3_0",
              "body_mass_index_bmi_f21001_0_0",
              "treat")
my_ukb_data <- subset(my_ukb_data, select=min_vars)
dim(my_ukb_data)
object.size(my_ukb_data)

# Remove all non essential variables
#keepers <- c("my_ukb_data")
#rm(list=setdiff(ls(), keepers))

# Extract minimal data frame w/ variables needed for matching (no NA)
match_vars <- c("eid",
              "age_years",
              "sex_f31_0_0",
              "treat")
min_data <- subset(my_ukb_data, select=match_vars)
dim(min_data)
object.size(min_data)
```

## Try MatchIt for Case Control Matching
```{r}
m.out <- matchit(treat ~ age_years + sex_f31_0_0, data = min_data, method="nearest", ratio=3)
```

## MatchIt output
```{r}
# Check balance
summary(m.out) 

# Extract Output data
matched_data <- match.data(m.out)
matched_cases <- which(matched_data$treat == 1)
matched_controls <- which(matched_data$treat == 0)

# Extract Indices in main data frame based on EID
matched_eids <- matched_data$eid
mapping <- sapply(matched_eids, function(x) which(my_ukb_data$eid == x))

# Test correctness
sum(my_ukb_data$eid[mapping] == matched_data$eid) == dim(matched_data)[1]
```

# Create TEXTFILE_OF_12.5K_IIDS.txt (3139 cases, 9417 controls)
```{r}
eids <- my_ukb_data$eid[mapping]
z <- cbind(eids,eids)
#fname <- "case_control_ids_matched.txt"
#fname <- "case_control_ids_cp_only.txt"
fname <- "case_control_ids_ap_only.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

# Create pancreatitis_vs_control_phenotypes.txt
```{r}
pheno <- my_ukb_data$treat[mapping]
z <- cbind(eids, eids, pheno)
#fname <- "pancreatitis_vs_control_phenotypes_matched.txt"
#fname <- "pancreatitis_vs_control_phenotypes_cp_only.txt"
fname <- "pancreatitis_vs_control_phenotypes_ap_only.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

```{r}
# Error check that phenotypes agree with pancreatitis cases
length(pheno)
length(eids)

a <- which(pheno == 1)
b <- which(pheno == 0)

length(a)
length(b)

a_id <- eids[a]
b_id <- eids[b]

a_id_sort <- sort(a_id)
b_id_sort <- sort(b_id)

intersect(a_id_sort,b_id_sort)

sum(setdiff(a_id_sort,b_id_sort) == a_id_sort)
sum(setdiff(b_id_sort,a_id_sort) == b_id_sort)

#main_id <- my_ukb_data$eid[panc_idc10_ix]
#main_id <- my_ukb_data$eid[cp_icd10_k86x]
main_id <- my_ukb_data$eid[ap_icd10_k85x]
main_id_sort <- sort(main_id)

sum(a_id_sort == main_id_sort)
sum(a_id_sort == main_id_sort) == length(a)
```

# Create pancreatitis_vs_control_discrete_covars.txt
```{r}

# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0[mapping]
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0[mapping]
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0[mapping]
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0[mapping]
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0[mapping]

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)
table(sex.c)
table(smoking.0c)
table(smoking.1c)
table(smoking.2c)
table(smoking.3c)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" or NA, then NA
smoking <- ifelse(
            (
              (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
              (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
              (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
              (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
            ),
            "1",
            ifelse(
              (
                (smoking.0c == "No") |
                (smoking.1c == "No") | 
                (smoking.2c == "No") | 
                (smoking.3c == "No")
              ),
              "0",
              "NA"
            )
)
z <- cbind(eids, eids, sex.c, smoking)

# Discover NA indices for error checking
na.ix <- which(is.na(z[,4]))

#fname <- "pancreatitis_vs_control_discrete_covars_matched.txt"
#fname <- "pancreatitis_vs_control_discrete_covars_cp_only.txt"
fname <- "pancreatitis_vs_control_discrete_covars_ap_only.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```

# Create pancreatitis_vs_control_quantitative_covars.txt 
```{r}
# Extract BMI, all other quantititative covars are PCs which will be added later

# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0[mapping]

# TODO(DMS): Alternately, can do an avg of intances 0-3??

# Look at BMI avgs
summary(bmi)

z <- cbind(eids, eids, bmi)
#fname <- "pancreatitis_vs_control_quantitative_covars_bmiOnly_matched.txt"
fname <- "pancreatitis_vs_control_quantitative_covars_bmiOnly_cp_only.txt"
#fname <- "pancreatitis_vs_control_quantitative_covars_bmiOnly_ap_only.txt"
write.table(z, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```


```{r}
#pca <- read.delim(file="ukb_matched.eigenvec", header=FALSE, sep=" ", dec=".")
#pca <- read.delim(file="cp_only.eigenvec", header=FALSE, sep=" ", dec=".")
pca <- read.delim(file="ukb_aponly.eigenvec", header=FALSE, sep=" ", dec=".")
pca$V2 <- NULL
#zeta <- as.data.frame(z)
#zeta <- read.delim(file="pancreatitis_vs_control_quantitative_covars_bmiOnly_matched.txt", header=FALSE, sep=" ", dec=".")
#zeta <- read.delim(file="pancreatitis_vs_control_quantitative_covars_bmiOnly_cp_only.txt", header=FALSE, sep=" ", dec=".")
zeta <- read.delim(file="pancreatitis_vs_control_quantitative_covars_bmiOnly_ap_only.txt", header=FALSE, sep=" ", dec=".")
zeta$ord <- 1:nrow(zeta) 
quant <- merge(zeta, pca, by.x="V1", by.y="V1", all.x=TRUE)
quant <- quant[order(quant$ord), ]
quant$ord <- NULL
quant <- as.matrix(quant)

#fname <- "pancreatitis_vs_control_quantitative_covars_matched.txt"
#fname <- "pancreatitis_vs_control_quantitative_covars_cp_only.txt"
fname <- "pancreatitis_vs_control_quantitative_covars_ap_only.txt"
write.table(quant, fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
```