---
  title: "gwas_input_phail"
output: html_document
---
  
  
RUN = *Cmd+Shift+Enter*.  
ADD CHUNK = *Cmd+Option+I*.  
PREVIEW = *Cmd+Shift+K*.    

## Install Packages/Libaries
```{r}
# Install packages
#devtools::install_github("kenhanscombe/ukbtools", dependencies = TRUE)

# Add libraries
library(ukbtools)
```

## Load up data into data frame
```{r}
# Create data frames for second data release
my_ukb_data <- ukb_df("ukb40721")
dim(my_ukb_data)
```

## Define variables to keep in data frame
```{r}
# Define all necessary disease columns for IDC10. Arrays 0-65.
col_disease_icd10 <- c("diagnoses_icd10_f41270_0_0","diagnoses_icd10_f41270_0_1",
                       "diagnoses_icd10_f41270_0_2","diagnoses_icd10_f41270_0_3",
                       "diagnoses_icd10_f41270_0_4","diagnoses_icd10_f41270_0_5",
                       "diagnoses_icd10_f41270_0_6","diagnoses_icd10_f41270_0_7",
                       "diagnoses_icd10_f41270_0_8","diagnoses_icd10_f41270_0_9",
                       "diagnoses_icd10_f41270_0_10","diagnoses_icd10_f41270_0_11",
                       "diagnoses_icd10_f41270_0_12","diagnoses_icd10_f41270_0_13",
                       "diagnoses_icd10_f41270_0_14","diagnoses_icd10_f41270_0_15",
                       "diagnoses_icd10_f41270_0_16","diagnoses_icd10_f41270_0_17",
                       "diagnoses_icd10_f41270_0_18","diagnoses_icd10_f41270_0_19",
                       "diagnoses_icd10_f41270_0_20","diagnoses_icd10_f41270_0_21",
                       "diagnoses_icd10_f41270_0_22","diagnoses_icd10_f41270_0_23",
                       "diagnoses_icd10_f41270_0_24","diagnoses_icd10_f41270_0_25",
                       "diagnoses_icd10_f41270_0_26","diagnoses_icd10_f41270_0_27",
                       "diagnoses_icd10_f41270_0_28","diagnoses_icd10_f41270_0_29",
                       "diagnoses_icd10_f41270_0_30","diagnoses_icd10_f41270_0_31",
                       "diagnoses_icd10_f41270_0_32","diagnoses_icd10_f41270_0_33",
                       "diagnoses_icd10_f41270_0_34","diagnoses_icd10_f41270_0_35",
                       "diagnoses_icd10_f41270_0_36","diagnoses_icd10_f41270_0_37",
                       "diagnoses_icd10_f41270_0_38","diagnoses_icd10_f41270_0_39",
                       "diagnoses_icd10_f41270_0_40","diagnoses_icd10_f41270_0_41",
                       "diagnoses_icd10_f41270_0_42","diagnoses_icd10_f41270_0_43",
                       "diagnoses_icd10_f41270_0_44","diagnoses_icd10_f41270_0_45",
                       "diagnoses_icd10_f41270_0_46","diagnoses_icd10_f41270_0_47",
                       "diagnoses_icd10_f41270_0_48","diagnoses_icd10_f41270_0_49",
                       "diagnoses_icd10_f41270_0_50","diagnoses_icd10_f41270_0_51",
                       "diagnoses_icd10_f41270_0_52","diagnoses_icd10_f41270_0_53",
                       "diagnoses_icd10_f41270_0_54","diagnoses_icd10_f41270_0_55",
                       "diagnoses_icd10_f41270_0_56","diagnoses_icd10_f41270_0_57",
                       "diagnoses_icd10_f41270_0_58","diagnoses_icd10_f41270_0_59",
                       "diagnoses_icd10_f41270_0_60","diagnoses_icd10_f41270_0_61",
                       "diagnoses_icd10_f41270_0_62","diagnoses_icd10_f41270_0_63",
                       "diagnoses_icd10_f41270_0_64","diagnoses_icd10_f41270_0_65",
                       "diagnoses_icd10_f41270_0_66","diagnoses_icd10_f41270_0_67",
                       "diagnoses_icd10_f41270_0_68","diagnoses_icd10_f41270_0_69",
                       "diagnoses_icd10_f41270_0_70","diagnoses_icd10_f41270_0_71",
                       "diagnoses_icd10_f41270_0_72","diagnoses_icd10_f41270_0_73",
                       "diagnoses_icd10_f41270_0_74","diagnoses_icd10_f41270_0_75",
                       "diagnoses_icd10_f41270_0_76","diagnoses_icd10_f41270_0_77",
                       "diagnoses_icd10_f41270_0_78","diagnoses_icd10_f41270_0_79",
                       "diagnoses_icd10_f41270_0_80","diagnoses_icd10_f41270_0_81",
                       "diagnoses_icd10_f41270_0_82","diagnoses_icd10_f41270_0_83",
                       "diagnoses_icd10_f41270_0_84","diagnoses_icd10_f41270_0_85",
                       "diagnoses_icd10_f41270_0_86","diagnoses_icd10_f41270_0_87",
                       "diagnoses_icd10_f41270_0_88","diagnoses_icd10_f41270_0_89",
                       "diagnoses_icd10_f41270_0_90","diagnoses_icd10_f41270_0_91",
                       "diagnoses_icd10_f41270_0_92","diagnoses_icd10_f41270_0_93",
                       "diagnoses_icd10_f41270_0_94","diagnoses_icd10_f41270_0_95",
                       "diagnoses_icd10_f41270_0_96","diagnoses_icd10_f41270_0_97",
                       "diagnoses_icd10_f41270_0_98","diagnoses_icd10_f41270_0_99",
                       "diagnoses_icd10_f41270_0_100","diagnoses_icd10_f41270_0_101",
                       "diagnoses_icd10_f41270_0_102","diagnoses_icd10_f41270_0_103",
                       "diagnoses_icd10_f41270_0_104","diagnoses_icd10_f41270_0_105",
                       "diagnoses_icd10_f41270_0_106","diagnoses_icd10_f41270_0_107",
                       "diagnoses_icd10_f41270_0_108","diagnoses_icd10_f41270_0_109",
                       "diagnoses_icd10_f41270_0_110","diagnoses_icd10_f41270_0_111",
                       "diagnoses_icd10_f41270_0_112","diagnoses_icd10_f41270_0_113",
                       "diagnoses_icd10_f41270_0_114","diagnoses_icd10_f41270_0_115",
                       "diagnoses_icd10_f41270_0_116","diagnoses_icd10_f41270_0_117",
                       "diagnoses_icd10_f41270_0_118","diagnoses_icd10_f41270_0_119",
                       "diagnoses_icd10_f41270_0_120","diagnoses_icd10_f41270_0_121",
                       "diagnoses_icd10_f41270_0_122","diagnoses_icd10_f41270_0_123",
                       "diagnoses_icd10_f41270_0_124","diagnoses_icd10_f41270_0_125",
                       "diagnoses_icd10_f41270_0_126","diagnoses_icd10_f41270_0_127",
                       "diagnoses_icd10_f41270_0_128","diagnoses_icd10_f41270_0_129",
                       "diagnoses_icd10_f41270_0_130","diagnoses_icd10_f41270_0_131",
                       "diagnoses_icd10_f41270_0_132","diagnoses_icd10_f41270_0_133",
                       "diagnoses_icd10_f41270_0_134","diagnoses_icd10_f41270_0_135",
                       "diagnoses_icd10_f41270_0_136","diagnoses_icd10_f41270_0_137",
                       "diagnoses_icd10_f41270_0_138","diagnoses_icd10_f41270_0_139",
                       "diagnoses_icd10_f41270_0_140","diagnoses_icd10_f41270_0_141",
                       "diagnoses_icd10_f41270_0_142","diagnoses_icd10_f41270_0_143",
                       "diagnoses_icd10_f41270_0_144","diagnoses_icd10_f41270_0_145",
                       "diagnoses_icd10_f41270_0_146","diagnoses_icd10_f41270_0_147",
                       "diagnoses_icd10_f41270_0_148","diagnoses_icd10_f41270_0_149",
                       "diagnoses_icd10_f41270_0_150","diagnoses_icd10_f41270_0_151",
                       "diagnoses_icd10_f41270_0_152","diagnoses_icd10_f41270_0_153",
                       "diagnoses_icd10_f41270_0_154","diagnoses_icd10_f41270_0_155",
                       "diagnoses_icd10_f41270_0_156","diagnoses_icd10_f41270_0_157",
                       "diagnoses_icd10_f41270_0_158","diagnoses_icd10_f41270_0_159",
                       "diagnoses_icd10_f41270_0_160","diagnoses_icd10_f41270_0_161",
                       "diagnoses_icd10_f41270_0_162","diagnoses_icd10_f41270_0_163",
                       "diagnoses_icd10_f41270_0_164","diagnoses_icd10_f41270_0_165",
                       "diagnoses_icd10_f41270_0_166","diagnoses_icd10_f41270_0_167",
                       "diagnoses_icd10_f41270_0_168","diagnoses_icd10_f41270_0_169",
                       "diagnoses_icd10_f41270_0_170","diagnoses_icd10_f41270_0_171",
                       "diagnoses_icd10_f41270_0_172","diagnoses_icd10_f41270_0_173",
                       "diagnoses_icd10_f41270_0_174","diagnoses_icd10_f41270_0_175",
                       "diagnoses_icd10_f41270_0_176","diagnoses_icd10_f41270_0_177",
                       "diagnoses_icd10_f41270_0_178","diagnoses_icd10_f41270_0_179",
                       "diagnoses_icd10_f41270_0_180","diagnoses_icd10_f41270_0_181",
                       "diagnoses_icd10_f41270_0_182","diagnoses_icd10_f41270_0_183",
                       "diagnoses_icd10_f41270_0_184","diagnoses_icd10_f41270_0_185",
                       "diagnoses_icd10_f41270_0_186","diagnoses_icd10_f41270_0_187",
                       "diagnoses_icd10_f41270_0_188","diagnoses_icd10_f41270_0_189",
                       "diagnoses_icd10_f41270_0_190","diagnoses_icd10_f41270_0_191",
                       "diagnoses_icd10_f41270_0_192","diagnoses_icd10_f41270_0_193",
                       "diagnoses_icd10_f41270_0_194","diagnoses_icd10_f41270_0_195",
                       "diagnoses_icd10_f41270_0_196","diagnoses_icd10_f41270_0_197",
                       "diagnoses_icd10_f41270_0_198","diagnoses_icd10_f41270_0_199",
                       "diagnoses_icd10_f41270_0_200","diagnoses_icd10_f41270_0_201",
                       "diagnoses_icd10_f41270_0_202","diagnoses_icd10_f41270_0_203",
                       "diagnoses_icd10_f41270_0_204","diagnoses_icd10_f41270_0_205",
                       "diagnoses_icd10_f41270_0_206","diagnoses_icd10_f41270_0_207",
                       "diagnoses_icd10_f41270_0_208","diagnoses_icd10_f41270_0_209",
                       "diagnoses_icd10_f41270_0_210","diagnoses_icd10_f41270_0_211",
                       "diagnoses_icd10_f41270_0_212")

# Additional variables needed for analysis
gwas_vars <- c("eid",
               "year_of_birth_f34_0_0","month_of_birth_f52_0_0",
               "sex_f31_0_0",
               "current_tobacco_smoking_f1239_0_0","current_tobacco_smoking_f1239_1_0",
               "current_tobacco_smoking_f1239_2_0","current_tobacco_smoking_f1239_3_0",
               "ethnic_background_f21000_0_0",
               "ethnic_background_f21000_1_0",
               "ethnic_background_f21000_2_0",
               "body_mass_index_bmi_f21001_0_0")
```

## Filter to keep only these needed variables in data frame
```{r}
my_ukb_data <- subset(my_ukb_data, select=c(col_disease_icd10,gwas_vars))
```

## Add in extra variables to dataframe as needed; age, etc. 
```{r}
# Add birthday variable, in R date format
my_ukb_data$birthday <- as.Date(sprintf("%d-%s-%d", my_ukb_data$year_of_birth_f34_0_0, 
                                        my_ukb_data$month_of_birth_f52_0_0,1), format="%Y-%B-%d")
# Add age
my_ukb_data$age_years <- as.numeric(floor((Sys.Date() - my_ukb_data$birthday)/365.25))
```

# Extract errant data point from data frame
```{r}
my_ukb_data <- my_ukb_data[-c(223169),]
```

## Define constants for downstream analysis
```{r}
# Create list of all possible indicies
num_subjects <- nrow(my_ukb_data)
num_var <- ncol(my_ukb_data)
all_indices <- seq(1, num_subjects, 1)
```

## Grab data frame indices for AP, CP, pancreatitis (AP & CP) via ICD10
```{r}
# AP ICD10
# Extract patients with K850-K859 acute pancreatitis designation via ICD 10
ap_conditions_icd10 <- c("K85","K850","K851","K852","K853", "K858", "K859")
ap_icd10_k85x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", ap_conditions_icd10))>0)
paste0("# with ICD10 AP ", length(ap_icd10_k85x))

# CP ICD10
# Extract patients with K860-861 chronic pancreatitis designation via ICD 10
cp_conditions_icd10 <- c("K860","K861")
cp_icd10_k86x <- which(rowSums(sapply(my_ukb_data[,col_disease_icd10],"%in%", cp_conditions_icd10))>0)
paste0("# with ICD10 CP ", length(cp_icd10_k86x))

# AP & CP ICD10
panc_idc10_ix <- sort(union(ap_icd10_k85x, cp_icd10_k86x))
paste0("# with ICD10 AP or CP ", length(panc_idc10_ix))
paste0("# with both ICD10 AP and CP ", length(intersect(ap_icd10_k85x, cp_icd10_k86x)))
```

## Grab data frame indices for controls based on random sampling of non AP/CP
```{r}
# CONTROLS ICD10
# Extract patients WITHOUT ICD10 AP or CP designation
control_icd10_total <- setdiff(all_indices, panc_idc10_ix)
paste0("# total possible controls ", length(control_icd10_total))

# Grab samples of control population
mult_factor <- 12 # Memory issues
s <- length(panc_idc10_ix) * mult_factor
set.seed(5750)
control_icd10_sample <- sample(control_icd10_total, s, replace = FALSE, prob = NULL)
paste0("# controls with 12:1 control:case ratio ", length(control_icd10_sample))

# Test for correctness
(nrow(my_ukb_data) - length(control_icd10_total)) == (length(ap_icd10_k85x) + 
                                                        length(cp_icd10_k86x) - length(intersect(ap_icd10_k85x,cp_icd10_k86x)))
```

## Grab data frame indices for controls from previous run
```{r}
control_fname <- "control_panc_allancestry_12x_eid.txt"
control_eids <- read.delim(file=control_fname, header=FALSE, sep="\t", dec=".")
control_eids <- as.array(control_eids[,1])
control_icd10_sample <- which(my_ukb_data$eid %in% control_eids)
```



## Identify indices for AP-only, CP(AP-CP inclusive), and related controls
```{r}
# Find indices
ap_cp <- intersect(ap_icd10_k85x, cp_icd10_k86x)
ap_only <- setdiff(ap_icd10_k85x, ap_cp)
cp_only <- setdiff(cp_icd10_k86x, ap_cp)

# AP-only Control
s <- length(ap_only) * mult_factor
set.seed(5750)
# could also sample from control_icd10_total if you didn't want to be a subset of control_icd10_sample
control_icd10_ap <- sample(control_icd10_sample, s, replace = FALSE, prob = NULL) 
paste0("# AP controls with 12:1 control:case ratio ", length(control_icd10_ap))
paste0("original # control ", length(control_icd10_sample))

# CP (AP-CP inclusive) Control
s <- length(cp_icd10_k86x) * mult_factor
set.seed(5750)
control_icd10_cp <- sample(control_icd10_sample, s, replace = FALSE, prob = NULL)
paste0("# CP controls with 12:1 control:case ratio ", length(control_icd10_cp))
```

##  Identify Caucasian samples and generate caucasian ancestry case/control indices
```{r}
# Write table
#table(my_ukb_data$ethnic_background_f21000_0_0)

# myukbdata indicies with caucasian ancestry
white <- with(my_ukb_data,
              which(ethnic_background_f21000_0_0 %in%
                      c("White","Any other white background","British",
                        "Irish")))
# print length of findings
dim(my_ukb_data)[1]
length(white)

# Filter data frame
#my_ukb_data <- my_ukb_data[white,]
#white <- my_ukb_data[white,]

# Indices
ap_only_caucasian <- intersect(ap_only, white)
cp_caucasian <- intersect(cp_icd10_k86x, white)
panc_caucasian <- intersect(panc_idc10_ix, white)

# Update control indices, grabbing from same sample population, appending extras if needed (no repetition)
control_caucasian <- intersect(control_icd10_sample, white)
control_caucasian_total <- intersect(setdiff(control_icd10_total,control_icd10_sample), white)
num_to_append <- (length(panc_caucasian) * mult_factor) - length(control_caucasian)
set.seed(5013)
append <- sample(control_caucasian_total, num_to_append, replace = FALSE, prob = NULL)
control_caucasian <- c(control_caucasian, append)

# Update control indices AP
s <- length(ap_only_caucasian) * mult_factor
set.seed(5750)
control_caucasian_ap <- sample(control_caucasian, s, replace = FALSE, prob = NULL)

# Update control indices CP
s <- length(cp_caucasian) * mult_factor
set.seed(5750)
control_caucasian_cp <- sample(control_caucasian, s, replace = FALSE, prob = NULL)
```

## Test that non-Caucasian samples are actually not caucasian
```{r}
# Testing for caucasian only indices
a <- setdiff(ap_only, white)
b <- setdiff(cp_icd10_k86x, white)
c <- setdiff(panc_idc10_ix, white)

for (ix in c) {
  test <- my_ukb_data[ix,]
  print(paste(test$ethnic_background_f21000_0_0))
}
```

## Generate outputs for all ancestries, all pancreatitis types
```{r}
# OUTPUT FILES FOR ALL ANCESTRIES, ALL PANCREATITIS TYPES
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_panc_allancestry_12x_eid.txt"
case_eid_fname <- "case_panc_allancestry_12x_eid.txt"
control_eid_fname <- "control_panc_allancestry_12x_eid.txt"
#pheno_fname <- "annotations_panc_allancestry_12x.txt"
pheno_fname <- "annotations_panc_allancestry_12x_ancestry.txt"

# case & control ix
ix_case <- panc_idc10_ix
ix_control <- control_icd10_sample

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]

#########################
# Ancestry (optional)
#########################
ancestry <- my_ukb_data$ethnic_background_f21000_0_0
levels(ancestry) <- c(
"other", #Prefer not to answer
"other", #Do not know
"white", #White
"mixed", #Mixed
"asian", #Asian or Asian British
"black", #Black or Black British
"chinese", #Chinese
"other", #Other ethnic group
"white", #British
"white", #Irish
"white", #Any other white background
"mixed", #White and Black Caribbean
"mixed", #White and Black African
"mixed", #White and Asian
"mixed", #Any other mixed background
"asian", #Indian
"asian", #Pakistani
"asian", #Bangladeshi
"asian", #Any other Asian background
"black", #Caribbean
"black", #African
"black" #Any other Black background
)
case_ancestry <- ancestry[ix_case]
control_ancestry <- ancestry[ix_control]

#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
# NOTE: ancestry coded as factor 1-6 (1=other,2=white,3=mixed,4=asian,5=black,6=chinese,NA)
#x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi, case_ancestry)
#y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi, control_ancestry)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR ALL ANCESTRIES, ALL PANCREATITIS TYPES
```

## Generate outputs for all ancestries, AP only
```{r}
# OUTPUT FILES FOR ALL ANCESTRIES, AP ONLY
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_ap_allancestry_12x_eid.txt"
case_eid_fname <- "case_ap_allancestry_12x_eid.txt"
control_eid_fname <- "control_ap_allancestry_12x_eid.txt"
pheno_fname <- "annotations_ap_allancestry_12x.txt"

# case & control ix
ix_case <- ap_only
ix_control <- control_icd10_ap

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]


#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR ALL ANCESTRIES, AP ONLY
```

## Generate outputs for all ancestries, CP(AP-CP inclusive)
```{r}
# OUTPUT FILES FOR ALL ANCESTRIES, CP (incl. AP-CP)
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_cp_allancestry_12x_eid.txt"
case_eid_fname <- "case_cp_allancestry_12x_eid.txt"
control_eid_fname <- "control_cp_allancestry_12x_eid.txt"
pheno_fname <- "annotations_cp_allancestry_12x.txt"

# case & control ix
ix_case <- cp_icd10_k86x
ix_control <- control_icd10_cp

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]


#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR ALL ANCESTRIES, CP (incl. AP-CP)
```

## Generate outputs for caucasian subjects, all pancreatitis types
```{r}
# OUTPUT FILES FOR CAUCASIAN SUBJECTS, ALL PANCREATITIS TYPES
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_panc_caucasian_12x_eid.txt"
case_eid_fname <- "case_panc_caucasian_12x_eid.txt"
control_eid_fname <- "control_panc_caucasian_12x_eid.txt"
pheno_fname <- "annotations_panc_caucasian_12x.txt"

# case & control ix
ix_case <- panc_caucasian
ix_control <- control_caucasian

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]


#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR CAUCASIAN SUBJECTS, ALL PANCREATITIS TYPES
```

## Generate outputs for caucasian subjects, AP only
```{r}
# OUTPUT FILES FOR CAUCASIAN SUBJECTS, AP ONLY
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_ap_caucasian_12x_eid.txt"
case_eid_fname <- "case_ap_caucasian_12x_eid.txt"
control_eid_fname <- "control_ap_caucasian_12x_eid.txt"
pheno_fname <- "annotations_ap_caucasian_12x.txt"

# case & control ix
ix_case <- ap_only_caucasian
ix_control <- control_caucasian_ap

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]


#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR CAUCASIAN SUBJECTS, AP ONLY
```

## Generate outputs for caucasian subjects, CP(AP-CP inclusive)
```{r}
# OUTPUT FILES FOR CAUCASIAN SUBJECTS, CP (incl. AP-CP)
#########################
#########################

# Names for 3 files of interest
casecontrol_eid_fname <- "casecontrol_cp_caucasian_12x_eid.txt"
case_eid_fname <- "case_cp_caucasian_12x_eid.txt"
control_eid_fname <- "control_cp_caucasian_12x_eid.txt"
pheno_fname <- "annotations_cp_caucasian_12x.txt"

# case & control ix
ix_case <- cp_caucasian
ix_control <- control_caucasian_cp

# GENERATE INDEX FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

# Extract eids
eids <- my_ukb_data$eid
x <- eids[ix_case]
y <- eids[ix_control]
z <- c(x,y)

# write to table
write.table(z, casecontrol_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(x, case_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
write.table(y, control_eid_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)

# GENERATE PHENOTYPE FILES FOR CASES/CONTROLS/BOTH
#########################
#########################

#########################
# EIDS
#########################
case_eids <-eids[ix_case]
control_eids <- eids[ix_control]

#########################
#SEX & SMOKING
#########################
# Extract sex and smoking
sex <- my_ukb_data$sex_f31_0_0
smoking.0 <-my_ukb_data$current_tobacco_smoking_f1239_0_0
smoking.1 <- my_ukb_data$current_tobacco_smoking_f1239_1_0
smoking.2 <- my_ukb_data$current_tobacco_smoking_f1239_2_0
smoking.3 <- my_ukb_data$current_tobacco_smoking_f1239_3_0

# Convert factors to characters
sex.c <- as.character(sex)
smoking.0c <- as.character(smoking.0)
smoking.1c <- as.character(smoking.1)
smoking.2c <- as.character(smoking.2)
smoking.3c <- as.character(smoking.3)

## Create master smoking vector
###############################
# If any of the instance index show smoking then 1, 
# if not, THEN if any of the indicies show NO smoking then 0,
# othewise, if all are "Prefer not to answer" then NA
smoking <- ifelse(
  (
    (smoking.0c %in% c("Yes, on most or all days", "Only occasionally")) | 
      (smoking.1c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.2c %in% c("Yes, on most or all days", "Only occasionally")) |
      (smoking.3c %in% c("Yes, on most or all days", "Only occasionally"))
  ),
  "1",
  ifelse(
    (
      (smoking.0c == "No") |
        (smoking.1c == "No") | 
        (smoking.2c == "No") | 
        (smoking.3c == "No")
    ),
    "0",
    "NA"
  )
)
case_sex <- sex.c[ix_case]
case_smoking <- smoking[ix_case]
control_sex <- sex.c[ix_control]
control_smoking <- smoking[ix_control]

#########################
# BMI
#########################
# Just using the first instance of BMI
bmi <- my_ukb_data$body_mass_index_bmi_f21001_0_0
case_bmi <- bmi[ix_case]
control_bmi <- bmi[ix_control]


#########################
# Output
#########################
x <- cbind(case_eids, 1, case_sex, case_smoking, case_bmi)
y <- cbind(control_eids, 0, control_sex, control_smoking, control_bmi)
z <- rbind(x,y)
write.table(z, pheno_fname, append=FALSE, sep=" ", row.names=FALSE, col.names=FALSE)
# END: OUTPUT FILES FOR CAUCASIAN SUBJECTS, CP (incl. AP-CP)
```

RUN = *Cmd+Shift+Enter*.  
ADD CHUNK = *Cmd+Option+I*.  
PREVIEW = *Cmd+Shift+K*.    

